Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down

### 2022-08-22 paper/ms.tex
Update ms.tex

\newcommand{\Nbody}{$N$-body}

@@

Our setup is initially out of equilibrium, but we found that after
$\sim500\,\textrm{Myr}$, the entire system has settled into a steady-state
configuration and initial transients appear not to affect the results after
this point. The constant surface density of the initial gas disk is important
for ensuring the gas disk is dense enough in order for comparisons to real
galaxies to be appropriate.

@@

We used a mass resolution of $7.5\times10^3\,M_{\odot}$ for the baryonic
components (initial stellar disk, stellar bulge, and gas) and a mass
resolution of $3.75\times10^4\,M_{\odot}$ for the dark matter halo. This mass
resolution is closest to ``level 3'' in the AURIGA
simulations\cite{2017MNRAS.467..179G}. This corresponds to approximately
$6.4\times10^6$ particles in the stellar disk, $1.1\times10^6$ in the bulge,
$1.2\times10^6$ in the gas disk, and $25.3\times10^6$ in the dark matter halo.
We used a softening length of $0.02\,\textrm{kpc}$ for all components.
Snapshots were saved at equal intervals of $0.005$ in the time units of the
simulation ($\textrm{kpc}/(\textrm{km}/\textrm{s})\sim1\,\textrm{Gyr}$).

>>>>

\newcommand{\Nbody}{$N$-body}
\newcommand{\AREPO}{\textsc{AREPO}}
\newcommand{\SMUGGLE}{SMUGGLE}

@@

\subsection{Numerical Model}
We use the Stars and MUltiphase Gas in GaLaxiEs (SMUGGLE) model
\citep{2019MNRAS.489.4233M} implemented within the moving-mesh, finite-volume
hydrodynamics and gravity code \AREPO{} \citep{2010MNRAS.401..791S}. The SMUGGLE
model additionally includes radiative heating and cooling, star formation, and
stellar feedback. Explicit gas cooling and heating of the multi-phase
interstellar medium is implemented, covering temperature ranges between $10$ and
$10^8\,\textrm{K}$.

Star formation occurs in cells above a density threshold
($n_{\textrm{th}}=100\,\textrm{cm}^{-3}$) with a star-formation efficiency of
$\epsilon = 0.01$. Star formation converts gas cells into star particles which
represent single stellar populations with a Chabrier initial mass
function \citep{2003PASP..115..763C}. For each star particle, the deposition of
energy, momentum, mass, and metals from stellar winds and supernovae is modeled.
Photo-ionization and radiation pressure are modeled using an approximate
treatment. A more detailed description of this model can be found in the
flagship SMUGGLE paper \citep{2019MNRAS.489.4233M}.

We used the fiducial model parameters, except that we increased the number of
effective neighbors $N_{\textrm{ngb}}$ for the deposition of feedback from
$64$ to $512$. We found that a lower value of $N_{\textrm{ngb}}$ resulted in
inefficient photo-ionization feedback since the photo-ionizing budget had not
been exhausted after deposition into $64$ neighboring cells. We also used an
updated version of SMUGGLE using a new mechanical feedback routine similar to
the one described in \citet{2018MNRAS.480..800H}. This updated routine is a
tensor renormalization which ensures linear and angular momentum conservation
to machine precision.

In addition to the SMUGGLE model, we considered a simpler model of the
interstellar medium based upon \citet{2003MNRAS.339..289S}. In this model, the
multiphase nature of the interstellar medium is modeled in a subgrid manner by
allowing each resolution element to have a ``cold'' and ``hot'' component, with
the equation of state of the gas suitably modified. Gas is allowed to
interchange between the ``cold'' and ``hot'' components through processes such
as cooling and stellar feedback. Cold gas is allowed to undergo star formation.
We refer to this model as the ``smooth interstellar medium'' model, and it is
described in more detail in \citet{2019MNRAS.489.4233M}.

\subsection{Bar Analysis}
The analysis of various bar properties is performed as follows. First, the
pattern speed is measured from the angle of the second Fourier component. We
measured the second Fourier component by computing,
\begin{equation}
\begin{split}
A_2 &= \sum_i m_i e^{i 2 \phi_i} \\
A_0 &= \sum_i m_i \textrm{,}
\end{split}
\end{equation}
where $m_i$ and $\phi_i$ are the mass and azimuthal angle of each particle,
respectively. We computed $A_2$ and $A_0$ in cylindrical bins of width
$0.5\,\textrm{kpc}$ from radii of $0$ to $30\,\textrm{kpc}$. We defined the
angle of the bar $\phi_b$ to be twice the angle of the complex number $A_2$ as
measured in the bin extending from a radius of $2.5$ to $3\,\textrm{kpc}$.
After correcting for the periodicity of $\phi_b$, we measured the pattern
speed as one-half the two-sided finite gradient of $\phi_b$ as a function of
time.

\begin{figure}
\centering
\includegraphics[width=9cm]{fig/A2_over_A0.pdf}
\caption{\textbf{Evolution of bar strength.} The bar strength is measured as
the maximum of the second Fourier component divided by the zeroth Fourier
component. We see that in the
\Nbody{} case (blue) the bar strength increases with time, consistent with
previous results showing that the bar strength increases as bars slow down. In
the SMUGGLE case (orange), we see that the bar strength remains roughly
constant, possibly slightly decreasing with time. This is also consistent with
the expected relation between pattern speed and strength since the bar in this
case is not slowing down.}
\label{fig:strength}
\end{figure}

\begin{figure*}
\centering
\includegraphics[width=18cm]{fig/bar_decomp.pdf}
\caption{\textbf{Disk decomposition into the barred and unbarred disk.} This
procedure is based on \citet{2016MNRAS.463.1952P} The \textit{left panel} shows
a face-on surface density projection through the stellar component of the
\SMUGGLE{} simulation (disk and bulge) at $t=1\,\textrm{Gyr}$. The
\textit{middle panel} shows the component of the disk identified as being
trapped in the bar while the \textit{right panel} shows the component of the
disk identified as not being trapped in the bar. The fact that the untrapped
stars form a roughly axisymmetric structure indicates our bar decomposition is
sufficiently accurate.}
\label{fig:decomp}
\end{figure*}

The time evolution of the bar strength, defined as the maximum of
$\left|A_2/A_0\right|$ as a function of radius, is shown in Extended Data
Fig.~\ref{fig:strength}. The quantity $\left|A_2/A_0\right|$ varies from $0$
to $1$, with larger values indicating a stronger bar pattern. We see that in
the \Nbody{} case, $\left|A_2/A_0\right|$ increases over time as the bar pattern
slows. This is consistent with previous \Nbody{} simulations which showed a
clear correlation between the bar pattern speed and the bar
strength\cite{2003MNRAS.341.1179A}. In the SMUGGLE case, we see that the bar
strength has an initial drop but then remains at a roughly constant, but
slightly decreasing, strength. This is consistent with the pattern speed in the
SMUGGLE case being roughly constant or slightly increasing.

Computing the length of the bar and the torque on the bar by different
components requires us to decompose the disk into a component which is trapped
by the bar and a component which is untrapped. In order to do this, we follow
closely the technique developed in ref.\cite{2016MNRAS.463.1952P} We analyzed
the orbit of each star particle (meaning initial disk, bulge, and newly formed
stars) by extracting the $x$-$y$ positions of the apoapse of each in a frame
corotating with the bar (apoapses are defined as local maxima in $r$). For
each apoapse, we searched for the $19$ closest apoapses in time and applied a
$k$-means clustering algorithm on this set of $20$ points with $k=2$. We then
computed for each of the two clusters the average angle from the bar
$\left<\Delta \phi\right>_{0,1}$, the standard deviation in $R$ of the points
${\sigma_R}_{0,1}$, and the average radius of the cluster
$\left<R\right>_{0,1}$. At each apoapse, a particle was considered to be in
the bar if it met the following criteria:
\begin{equation}
\textrm{max}\left(\left<\Delta \phi\right>_{0,1}\right) < \pi / 8
\end{equation}
\begin{equation}
\frac{{\sigma_R}_0 + {\sigma_R}_1}{\left<R\right>_0 + \left<R\right>_1} < 0.22
\end{equation}
These criterion are slightly different and simplified from the ones used in
ref.\cite{2016MNRAS.463.1952P}, but we found them to empirically work well at
decomposing the disk into a bar and disk component. In Extended Data
Fig.~\ref{fig:decomp}, we show an example of this decomposition. The
\textit{left} panel shows a surface density projection of the stellar disk and
bulge (including newly formed stars) from the SMUGGLE model after
$1\,\text{Gyr}$ of evolution in a frame such that the bar is aligned with the
$x$-axis. The \textit{middle} panel shows a projection of the subset of stars
that are identified as being trapped in the bar and the \textit{right} panel
shows a projection of the stars that are not identified as being trapped. The
fact that the \textit{right} panel is roughly axisymmetric indicates the bar
decomposition is performing adequately.

After the disk has been decomposed into a trapped and untrapped component, we
measured the bar length as being the radius $R_b$ which encapsulates $99\%$ of
the stars identified as being trapped in the bar, allowing for some outliers.
For the computation of the torque, we used the tree algorithm in
\texttt{MakeNewDisk}\cite{2005MNRAS.361..776S} customized to be accessible
from \texttt{Python} using \texttt{Cython}. This algorithm is based on the
\texttt{TREESPH} code\cite{1989ApJS...70..419H}. We constructed a tree with an opening angle of $0.35$ using
only the star particles identified as being trapped in the bar. We then queried the tree at the locations of all
resolution elements in the other components and computed the torque of the bar
on such components. The torque on the bar by the other components is simply
the negative of the torque on the other components by the bar. A similar
method was applied in measuring the torque for the disk when the pattern speed
is kept constant, as is done in Fig.~\ref{fig:equil}.

% \subsection{Idealized Mock Light Images}
% We created synthetic mock light images at $1024^2$ resolution in
% Fig.~\ref{fig:overview} for the \textit{Hubble Space Telescope} using a
% post-processed Monte Carlo radiation transfer simulation with the SKIRT9 code
% including secondary emission from dust\cite{2020AC....3100381C}. Emission from
% star particles was calculated assuming a Bruzual-Charlot spectral energy
% distribution\cite{2003MNRAS.344.1000B} and a Chabrier initial stellar mass
% function\cite{2003PASP..115..763C}. The luminosity from each star particle was
% smoothed using a cubic spline kernel with each star particle being assigned a
% smoothing length equal to its distance to its $32$nd closest neighbor. We
% launched $10^9$ photon packets per simulation segment. For star particles
% which were present in the initial simulation (i.e., initial disk and bulge
% particles), we assumed that they have an age of $5\,\textrm{Gyr}$ and solar
% metallicity. For star particles in the SMUGGLE simulation which are newly
% formed, we used their recorded ages and metallicities.

% For the SMUGGLE simulations, we included absorption and emission from dust. We
% assumed the THEMIS (The Heterogeneous dust Evolution Model for Interstellar
% Solids) dust model\cite{2017AA...602A..46J} to convert gas metallicity and
% mass density to dust mass density. For the \Nbody{} simulations we ran the
% SKIRT9 code in a mode without an obscuring medium, but otherwise applied the
% same procedure.

% We used the profiles for filters F814W, F606W, and F475W on the Advanced
% Camera for Surveys (ACS) instrument for red, green, and blue pixel values,
% respectively. Filtered fluxes were converted to RGB values using an $\arcsinh$
% scaling following ref.\cite{2004PASP..116..133L} Scalings and cutoff values
% were adjusted manually to create visually appealing images. Filter profiles
% were provided by the Spanish Virtual Observatory Filter
% Service\cite{2012ivoa.rept.1015R, 2020sea..confE.182R}.

\subsection{Plotting Details}
We saved snapshots in intervals of $0.005$ in the time units of the
simulation, $\textrm{kpc}/(\textrm{km}/\textrm{s})$, which is very nearly
equal to $1\,\textrm{Gyr}$ (it is $\sim0.977\,\textrm{Gyr}$). Therefore,
throughout this work we referred to the native code time unit as
$\textrm{Gyr}$. None of our results are sensitive to this choice.

In order to remove numerical noise in several quantities we computed, we
applied a Savitzky-Golay filter \citep{1964AnaCh..36.1627S} as implemented in
\texttt{scipy} using a window length of $81$ and polynomial order of $3$. This
filter was applied to plots of pattern speeds, bar lengths and strengths,
torques, and angle differences.

\section{Discussion}
Our setup is initially out of equilibrium, but we found that after
$\sim500\,\textrm{Myr}$, the entire system has settled into a steady-state
configuration and initial transients appear not to affect the results after
this point. The constant surface density of the initial gas disk is important
for ensuring the gas disk is dense enough in order for comparisons to real
galaxies to be appropriate.

@@

The pattern speed evolution of this setup is given in Extended Data
Fig.~\ref{fig:GFM}, which shows that the pattern speed evolves in a similar
manner to the SMUGGLE case shown in Fig.~\ref{fig:prop}.
